---
title: " `brms` workshop: compendium"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

```{r}
library(tidyverse)
library(brms)
library(tidybayes)
```

This document contains the code used during the workshop, so you can play along, copy some code, etc... 

The order is according to the slide show.

## Likelihood

We refered to 10 observations on people trying to finish 10k running (expressed in minutes). Let's create the dataset!

```{r}
## RT stands for Running Time

RT <- c(
  52,
  54,
  58,
  48,
  41, ## whoow; flying!
  49,
  72,
  53,
  64,
  62)
```

Calculate our 'maximum likelihood estimates' if we assume that running times are normally distributed with a mean and a standard deviation that define our probability density function.

```{r }
Mean_RT <- mean(RT, na.rm = T)
Mean_RT
Sd_RT   <- sd(RT, na.rm = T)
Sd_RT
```

Our 'model' for RT:

```{r}
X <- seq(30, 90, by = 0.5)

Y <- dnorm(X, 
           mean = Mean_RT, 
           sd = Sd_RT)

data_frame(X, Y) %>%
  ggplot(aes(x= X, y = Y)) + geom_line()

```

##  Prior elicitation

```{r}
## For the mean

c()

## For the sd


```



```{r}
Data <- data_frame(RT)

MA <- brm(RT~1,
          data = Data,
          backend = "cmdstanr"
          )

summary(MA)
```

