---
title: "Prerequisites for the workshop on Bayesian Analyses"
author: "Sven De Maeyer"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# Intro

For the workshop Bayesian analysis on 13 and 15 September we will be using R. To be able to participate properly, it is advisable to install a number of packages on your own PC beforehand, as well as to become familiar with the way I usually code in R. This document is a brief guide to help you prepare.

# Necessary Software & Packages

## CmdStan & CmdStanR

Bayesian estimation and analyses require some dedicated samplers that should be installed on your machine. The workflow that we will cover in the workshop relies on `Stan` software. `Stan` stands alone from R but can be called through R making use of two different ways. We will cover this more in the workshop. 

For the workshop I will use the `CmdStan` chain and the package `cmdstanr`. Therefore I would encourage you to download and install a working `cmdstan` and `cmdstanr`. The following vignette can help you accomplish this: https://mc-stan.org/cmdstanr/articles/cmdstanr.html 

Normally if you follow the steps described in that vignette you will succeed in installing and testing your installation. 

## `brms`

For the analyses we will rely on the amazing package `brms`. Installing this package is not that difficult, but maybe install before the workshop and start some exploring ;-)

## `tidyverse`

When I code I use the <b> functional programming </b> approach and a lot of the <b>tidy</b> principles (see below). The package `tidyverse` bundles a set of packages that are needed for this approach to coding. If you want to learn more on the whole universe of `tidyverse` you can explore the book of Hadley Wickham: https://r4ds.had.co.nz/

## `ggplot2`

Part of the `tidyverse` package is the `ggplot2` package. So it will be installed if you also install `tidyverse` normally. Nevertheless I find it worth mentioning separately because I think knowing how to make graphs with `ggplot2` is very handy even if you do not dive into bayesian analyses. A great resource to learn to visualize your data is the tutorial written by Cédric Scherer: https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/ !

He also blogs on creating great visualizations and has some nice talks that I think you can find on YouTube!

## `patchwork`

I mention this package separately as well. `patchwork` allows you to combine different plots created with `ggplot2`. This package should be installed on it's own.

## `tidybayes`

This package is a dedicated package to make use of the <b>tidy</b> principles when applying it to (results of) Bayesian analyses.

## `bayesplot`

We will use visualisations a lot in Bayesian analyses to summarize the information of models and parameter estimation. `bayesplot` bundles a number of functions that make this visualizing much easier (although it uses `ggplot2` under the hood so we could accomplish similar results just knowing `ggplot2`).

# Coding in the tidy universe...

There are a great number of roads that will lead to Rome! Coding in R can be accomplished in several ways. This means that code can get messy as well! Therefore, I try to keep myself as much as possible to some rules that are described by what is called the <b>tidy</b> way of coding. Some basic things are key to this approach. I will shortly describe them here so you will recognize this way of working when we're in the workshop. 

## The pipe

When coding I make use of what is called <b>the pipe operator</b> (notice that you can only use this when you loaded the `tidyverse` package).

The pipe in R code is the following: ```%>%```

What is nice about using the pipe operator is that you can read a piece of code from left to right focussing on the verbs (as Hadley Wickham will call it sometimes).

A short example. 

Imagine you have a vector of numbers for which you want to calculate a mean: ```c(1,2,3,4,5)```.

There are different ways to do this in R. 

A first one is doing it in separate lines of code:

```{r}
x <- c(1,2,3,4,5)
mean(x)
```

For this easy example this would work fine. But a typical side-effect of working this way is that we create a lot of objects along the way. Here we created the object `x` that stored the values of the vector.

To avoid this we could combine all the code in one line by embedding functions:

```{r}
mean(c(1,2,3,4,5))
```

One line of code and no storage of a new object. But this can get complicated if you want to combine a set of functions resulting in almost non-readable code! 

Then there is the pipe operator:

```{r}
library(tidyverse)
c(1,2,3,4,5) %>%
  mean()
```

This is how to read this code:

```{r}
c(1,2,3,4,5) %>% # create the vector and then
  mean()         # calculate the mean
```

So the pipe can be read as `take (or create) A` and `then do B` with the result. We could create a whole pipeline of functions to be applied in this way making use of the pipe operator (this becomes more clear in the next section when we shortly describe the `dplyr` verbs).

A tweet that I encountered gives a great analogy. Using the pipe you could write:

> I %>% woke up %>% took a shower %>% got breakfast %>% took the metro %>% arrived at work %>% ...

## `dplyr` verbs

Within the `tidyverse` universe we can make use of a great package called `dplyr` for a broad spectrum of data-management functions. The functions provided in that package are also sometimes called <b>verbs</b>:

- `mutate( )` to do recoding etc;
- `select( )` to select a subset of columns/variables;
- `filter( )` to filter cases;
- `group_by( )` to apply what's coming afterwards for specific groups in the data by a grouping variable;
- `arrange( )` to sort the data;
- `summarize( )` to apply summarizing functions (like taking the mean etc.);
- `rename( )` to rename columns;

These functions can be combined in a single statement making use of the pipe operator. 

Here are some examples on the use of these verbs applied to the built-in dataset `starwars` that is part of the `tidyverse` package.

Calculate the mean `birth_year` for all the characters:

```{r}
starwars %>% 
  summarize(
    mean_birth_year = mean(birth_year, na.rm = TRUE)
    ) %>%
  print()
```

Calculate the average height and the average mass by gender:

```{r}
starwars %>%
  group_by(gender) %>%
  summarize(
    mean_height = mean(height, na.rm = T),
    mean_mass = mean(mass, na.rm = T)
  ) 
```

Now, let's use the verbs to create a graph. We want to create a scatterplot with mass on the x-axis and height on the y-axis. Also, we want to add a third variable in the mix: whether the character's birth year is hither than 50 or not (just for fun). Finally we only want to show characters that have a mass lower than 200 and with a birth_year that is known.

```{r}
starwars %>%
  # First filter the cases with a mass higher than 200 and for which birth_year is not NA
  filter(
    mass < 200,
    !is.na(birth_year)
  ) %>%

  # Then we create a variable that has value yes if the character's birth year is above 50 and no otherwise
  mutate(
    birth_above_50 = if_else(birth_year > 50, "yes", "no") 
  ) %>%
  
  # Select the necessary variables for the graph (not really necessary but this is to show this option)
  select(
    height, 
    mass, 
    birth_above_50) %>%

  # Create the graph
  ggplot(
    aes(x = mass, y = height, color = birth_above_50)
  ) + geom_point()
```

